My title

### 2024-03-12 paper/paper.qmd
updated model and data writing

@@

```{r}
library(arrow)
library(dplyr)
survey_analysis_data <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/survey_analysis_data.parquet")
# poststrat_analysis_data <- read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/poststrat_analysis_data.parquet")
print(survey_analysis_data)

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-survey_presidential_preferences
#| fig-cap: The Distribution of Presidential Preferences, by gender and highest level of education
survey_analysis_data |>
  ggplot(aes(x = education_level, fill = vote24)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(sex)) +
  theme_minimal() +
  labs(
    x = "Highest education",
    y = "Number of respondents",
    fill = "Preferred Presidential Candidate"
  ) +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

## Survey Data {#sec-survey}

@@

```{r}
#| label: fig-survey_presidential_preferences1
#| fig-cap: The Distribution of Presidential Preferences, by gender and race
#| echo: false

@@

```{r}
#| label: fig-electoral-map-survey
#| fig-cap: Electoral map based solely on the survey data
#| echo: false
library(ggplot2)
library(statebins)
survey_electoral_map <- survey_analysis_data %>% 
  group_by(state, vote_biden) %>% 
  summarise(n = n()) %>% 
  group_by(state) %>% 
  mutate(prop = n/sum(n)) %>% 
  filter(vote_biden == 1)

@@

@@

@citeIPUMS

@@

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

@@

# poststrat_analy <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/poststrat_analysis_data.parquet")
# 
# poststrat_data_cells <-
#   poststrat_analysis_data |>
#   count(state, sex, age_group, education_level)
# 
# poststrat_data_cells <-
#   poststrat_data_cells |>
#   mutate(prop = n / sum(n),
#          .by = stateicp)
# 
# poststrat_data_cells

@@

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

The logistic regression used to predict ... is as follows:

```{=tex}
\begin{align}
log(\frac{p}{1-p}) = \beta_0 + \beta_1 \times x_{sex} + \beta_2 \times x_{age\_bracket} +  \beta_3 \times x_{races} + \\
\beta_3 \times x_{education\_level} + \beta_4 \times x_{state} + \beta_5 
\times x_{urban} + \beta_5 \times x_{race\_black}
\end{align}
```
We run our model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

@@

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

>>>>

Packages used for data cleaning, model, visualizations, and tables: `tidyverse` @citetidyverse, `janitor` @citejanitor, `rstanarm`, `readr`, `ggplot2`, `knitr`, and `arrow`.

@@

library(arrow)
library(dplyr)
survey_analysis_data <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/survey_analysis_data.parquet")
# poststrat_analysis_data <- read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/poststrat_analysis_data.parquet")
print(survey_analysis_data)

@@

```{r}
#| echo: false
#| warning: false
#| label: fig-survey_presidential_preferences
#| fig-cap: The Distribution of Presidential Preferences, by gender and highest level of education
survey_analysis_data |>
  ggplot(aes(x = education_level, fill = vote24)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(sex)) +
  theme_minimal() +
  labs(
    x = "Highest education",
    y = "Number of respondents",
    fill = "Preferred Presidential Candidate"
  ) +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

```{r}

## Individual-Level Survey Data {#sec-survey}

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-survey_presidential_preferences1
#| fig-cap: The Distribution of Presidential Preferences, by gender and race

@@

```{r}
# write about America's urban-rural divide
#| echo: false
#| warning: false
#| label: fig-survey_urban_rural
#| fig-cap: The Distribution of Presidential Preferences of Respondents Living in Urban vs Rural Areas
survey_analysis_data |>
  ggplot(aes(x = urban, fill = vote24)) +
  stat_count(position = "dodge") +
  # facet_wrap(facets = vars(sex)) +
  theme_minimal() +
  labs(
    y = "Number of respondents",
    fill = "Preferred Presidential Candidate"
  ) +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

# do a table of the proportion of respondents for trump vs biden in each state
# overall is like 51% biden
```

```{r}
#| echo: false
#| warning: false
#| label: tbl-survey_presidential_preferences_table
#| fig-cap: Who Voters Plan to Support (Survey Data)

num_respondents <- nrow(survey_analysis_data)
num_biden <- sum(survey_analysis_data$vote_biden)
num_trump <- num_respondents - num_biden

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-electoral-map-survey
#| fig-cap: Electoral map based solely on the survey data
library(ggplot2)
library(statebins)
survey_electoral_map <- survey_analysis_data %>% 
  group_by(state, vote_biden) %>% 
  summarise(n = n()) %>% 
  group_by(state) %>% 
  mutate(prop = n/sum(n)) %>% 
  filter(vote_biden == 1)

@@

@fig-electoral-map-survey illustrates the proportion of voters in each U.S. state who plan to support President Biden in the 2024 Presidential Election. We can see that there is strong support for President Biden in Washington, Oregon, New York, New Jersey, Maryland, Massachusetts, and Connecticut. We also see that Oklahoma, Idaho, and Utah heavily favor former President Trump. These results are expected and there were similar margins in the 2020 election. These states have also been Dem and GOP strongholds since 1992 (or whatever year). However, this electoral map also highlights some of the limitations fo teh survey data. Wyoming and Kansas are very blue in the electoral map based solely on the survey data, however that is very far from recent election results in those states. In fact, Wyoming was the most Pro-Trump state in both the 2016 and 2020 presidential elections, with Trump receiving 2016% and 2020% of the vote in each of those elections, respectively. The survey data set includes very few examples from small states.

@@

The post-stratification data used is comes from the American Community Survey (ACS) 2022 provided by IPUMS ( @citeIPUMS ).

@@

<!-- # Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details]. -->

@@

# poststrat_analy <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/poststrat_analysis_data.parquet")
# 
# poststrat_data_cells <-
#   poststrat_analysis_data |>
#   count(state, sex, age_group, education_level)
# 
# poststrat_data_cells <-
#   poststrat_data_cells |>
#   mutate(prop = n / sum(n),
#          .by = stateicp)
# 
# poststrat_data_cells
```

@@

Define $y_i$ whether an individual's preferred 2024 presidential candidate is Joe Biden (1) or Donald Trump (0). The logistic regression model used to predict support for Biden using sex, age, race, highest level of education, state, and whether the individual lives in an urban or rural area is as follows:

```{=tex}
\begin{align}
Pr(vote\_biden_i =1) = logit^{-1}( \beta_0 + \beta_1 \times x_{sexi} + \beta_2 \times x_{age\_bracketi} +  \beta_3 \times x_{racesi} + \\
\beta_3 \times x_{education\_leveli} + \beta_4 \times x_{statei} + \beta_5 
\times x_{urbani} + \beta_5 \times x_{race\_blacki})
\end{align}
```
We run our model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

@@

We expect to see increased support for President Biden amongst Black college-educated women in urban areas.
